Exploring mathematical objects from custom-tailored mathematical universes

### 2020-06-30 paper-filmat.tex
Explain subsets in Eff better (in response to review)

The usual proof sets up the halting function~$h : \NN \to \NN$, defined using
the case distinction
\[ h : n \mapsto \begin{cases}
  1, & \text{if the $n$-th Turing machine terminates}, \\
  0, & \text{if the $n$-th Turing machine does not terminate},
\end{cases} \]
and proceeds to show that~$h$ is not computable. However, in the \effective
topos, this definition does not give rise to a total function from~$\NN$
to~$\NN$. The actual domain is the subset of those natural numbers~$n$
for which the~$n$-th Turing machine
terminates or does not terminate. This condition is trivial only assuming the
law of excluded middle; intuitionistically, this condition is nontrivial and
cuts out a nontrivial subset of~$\NN$.

XXX: subsets...
\end{ex}

>>>>

The usual proof sets up the halting function~$h : \NN \to \NN$, defined using
the case distinction
\[ h : n \mapsto \begin{cases}
  1, & \text{if the $n$-th Turing machine terminates}, \\
  0, & \text{if the $n$-th Turing machine does not terminate},
\end{cases} \]
and proceeds to show that~$h$ is not computable. However, in the \effective
topos, this definition does not give rise to a total function from~$\NN$
to~$\NN$. The actual domain is the subset~$M$ of those natural numbers~$n$
for which the~$n$-th Turing machine
terminates or does not terminate. This condition is trivial only assuming the
law of excluded middle; intuitionistically, this condition is nontrivial and
cuts out a nontrivial subset of~$\NN$.

Subobjects in the \effective topos are more than mere subsets; to give an
element of~$M$ in the \effective topos, we need not only give a natural
number~$n$ such that the~$n$-th Turing machine terminates or does not
terminate, but also a computational witness of either case. For any particular
numeral~$n_0 \in \NN$, there is such a witness (appealing to the law of
excluded middle on the meta level), and hence the statement~``$n_0
\in M$'' holds in the \effective topos. However, there is no program which
could compute such witnesses for any number~$n$, hence the statement~``$\forall
n\?\NN\_ n \in M$'' is not true in~$\Eff$ and hence the \effective topos does not believe~$M$
and~$\NN$ to be the same.
\end{ex}
